---
title: "Hands-on Exercise 4b"
subtitle: "Lesson 4b: [Visual Statistical Analysis](https://r4va.netlify.app/chap10)" 
author: "Victoria Neo"
date: 02/6/2024
date-modified: last-modified
format:
  html:
    code-fold: true
    code-summary: "code block"
    code-tools: true
    code-copy: true
execute:
  warning: false
---

![*Taken from* [How to Plot Statistical Group Comparisons in R: ggstatsplot](https://www.youtube.com/watch?v=Yhz0TKP_CUc)](images/clipboard-1237587139.png){fig-align="left" width="356"}

# Overview Summary

|                  |                                                                                                                                                                                                    |
|------------------|----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------|
| Work done        | Hands-on Exercise 4b                                                                                                                                                                               |
| Hours taken      | ⏱️⏱️ (hospitalisation leave)                                                                                                                                                                       |
| Questions        | 0                                                                                                                                                                                                  |
| How do I feel?   | 😣                                                                                                                                                                                                 |
| What do I think? | I feel genuinely lost but this is an important topic to help improve the data visualisations with statistical tests (that was something I struggled to show in Take-home_Ex01 and Take-home_Ex02). |

# 1 Overview Notes

[**ggstatsplot**](https://indrajeetpatil.github.io/ggstatsplot/) is an extension of [**ggplot2**](https://ggplot2.tidyverse.org/) package for creating graphics with details from statistical tests included in the information-rich plots themselves.

-   To provide alternative statistical inference methods by default.

-   To follow best practices for statistical reporting. For all statistical tests reported in the plots, the default template abides by the [APA](https://my.ilstu.edu/~jhkahn/apastats.html) gold standard for statistical reporting.

# 2 Exam_data

## 2.1 Getting Started

### 2.1.1 Installing and loading the required libraries

The code chunk below uses p_load() of pacman package to check if the following R packages are installed in the computer. If they are, then they will be launched into R.

-   tidyverse, a family of R packages for data science process,

-   ggstatsplot, an extension of ggplot2 to create graphics with details from statistical tests included in the information-rich plots themselves.

```{r}
pacman::p_load(ggstatsplot, tidyverse)
```

### 2.1.2 Data Set

This section is taken from [Hands-on_Ex02](Hands-on_Ex/Hands-on_Ex01/Hands-on_Ex01.html) as we are using the same data set.

The data set, *Exam_data.csv,* contains the Year-end examination grades of a cohort of primary 3 students from a local school, and is uploaded as **exam_data**.

#### 2.1.2.1 Importing exam_data

In the code chunk below, read_csv() of readr package is used to import Exam_data.csv data file into R and save it as an tibble data frame called exam_data.

```{r}
exam <- read_csv("data/Exam_data.csv")

exam

```

## 2.2 Statistical Testing

### 2.2.1 **One-sample test: `gghistostats()` method**

In the code chunk below, `gghistostats()` is used to to build an visual of one-sample test on English scores.

::: panel-tabset
## The plot

```{r}
#| echo: false 
#| warning: false
set.seed(1234)

gghistostats(
  data = exam,
  x = ENGLISH,
  type = "bayes",
  test.value = 60,
  xlab = "English scores"
)
```

## The code chunk

```{r}
#| warning: false 
#| code-fold: show
#| eval: false
set.seed(1234)

gghistostats(
  data = exam,
  x = ENGLISH,
  type = "bayes",
  test.value = 60,
  xlab = "English scores"
)     
```
:::

#### 2.2.1.1 **Unpacking the Bayes Factor**

-   A Bayes factor is the ratio of the likelihood of one particular hypothesis to the likelihood of another. It can be interpreted as a measure of the strength of evidence in favor of one theory among two competing theories.

-   That’s because the Bayes factor gives us a way to evaluate the data in favor of a null hypothesis, and to use external information to do so. It tells us what the weight of the evidence is in favor of a given hypothesis.

-   When we are comparing two hypotheses, H1 (the alternate hypothesis) and H0 (the null hypothesis), the Bayes Factor is often written as B10.

-   The [**Schwarz criterion**](https://www.statisticshowto.com/bayesian-information-criterion/) is one of the easiest ways to calculate rough approximation of the Bayes Factor.

-   Bayes factors have been offered by Bayesians as alternatives to P values (or significance probabilities) for testing hypotheses and for quantifying the degree to which observed data support or conflict with a hypothesis.

#### 2.2.1.2 **How to interpret Bayes Factor**

A **Bayes Factor** can be any positive number. What the Bayes factor actually measures is the change in the odds in favor of the hypothesis when going from the prior to the posterior.

### 2.2.2 **Two-sample mean test: `ggbetweenstats()`**

In the code chunk below, `ggbetweenstats()` is used to build a visual for two-sample mean test of Maths scores by gender.

::: panel-tabset
## The plot

```{r}
#| echo: false 
#| warning: false
ggbetweenstats(
  data = exam,
  x = GENDER, 
  y = MATHS,
  type = "np",
  messages = FALSE
)
```

## The code chunk

```{r}
#| warning: false 
#| code-fold: show
#| eval: false
ggbetweenstats(
  data = exam,
  x = GENDER, 
  y = MATHS,
  type = "np",
  messages = FALSE
) 
```
:::

### 2.2.3 **Oneway ANOVA Test: `ggbetweenstats()` method**

In the code chunk below, `ggbetweenstats()` is used to build a visual for One-way ANOVA test on English score by race.

-   “ns” → only non-significant

-   “s” → only significant

-   “all” → everything

::: panel-tabset
## The plot

```{r}
#| echo: false 
#| warning: false
ggbetweenstats(
  data = exam,
  x = RACE, 
  y = ENGLISH,
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "s",
  p.adjust.method = "fdr",
  messages = FALSE
)
```

## The code chunk

```{r}
#| warning: false 
#| code-fold: show
#| eval: false
ggbetweenstats(
  data = exam,
  x = RACE, 
  y = ENGLISH,
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "s",
  p.adjust.method = "fdr",
  messages = FALSE
)
```
:::

### 2.2.4 **Significant Test of Correlation: `ggscatterstats()`**

In the code chunk below, `ggscatterstats()` is used to build a visual for Significant Test of Correlation between Maths scores and English scores.

::: panel-tabset
## The plot

```{r}
#| echo: false 
#| warning: false
ggscatterstats(
  data = exam,
  x = MATHS,
  y = ENGLISH,
  marginal = FALSE,
  )
```

## The code chunk

```{r}
#| warning: false 
#| code-fold: show
#| eval: false
ggscatterstats(
  data = exam,
  x = MATHS,
  y = ENGLISH,
  marginal = FALSE,
  )
```
:::

### 2.2.5 **Significant Test of Association (Dependence) : `ggbarstats()` methods**

In the code chunk below, the Maths scores is binned into a 4-class variable by using [*cut()*](https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/cut).

```{r}
exam1 <- exam %>% 
  mutate(MATHS_bins = 
           cut(MATHS, 
               breaks = c(0,60,75,85,100))
)
```

In this code chunk below [*ggbarstats()*](https://indrajeetpatil.github.io/ggstatsplot/reference/ggbarstats.html) is used to build a visual for Significant Test of Association.

::: panel-tabset
## The plot

```{r}
#| echo: false 
#| warning: false
ggbarstats(exam1, 
           x = MATHS_bins, 
           y = GENDER)
```

## The code chunk

```{r}
#| warning: false 
#| code-fold: show
#| eval: false
ggbarstats(exam1, 
           x = MATHS_bins, 
           y = GENDER)
```
:::

# 3 car_resale

## 3.1 Getting Started

In this section, you will learn how to visualise model diagnostic and model parameters by using parameters package.

-   Toyota Corolla case study will be used. The purpose of study is to build a model to discover factors affecting prices of used-cars by taking into consideration a set of explanatory variables. Please see above \[2.2 car_resale\].

### 3.1.1 Installing and loading the required libraries

The code chunk below uses p_load() of pacman package to check if the following R packages are installed in the computer. If they are, then they will be launched into R.

-   readxl, read_xls() of readxl package is used to import the Excel worksheet,

-   [performance](https://easystats.github.io/performance/), that provides utilities for computing indices of model quality and goodness of fit. These include measures like r-squared (R2), root mean squared error (RMSE) or intraclass correlation coefficient (ICC) , but also functions to check (mixed) models for overdispersion, zero-inflation, convergence or singularity,

-   [parameters](https://www.rdocumentation.org/packages/parameters/versions/0.21.4), contains utilities for processing the parameters of various statistical models, and

-   [see](https://easystats.github.io/see/), helps users to utilize visualisation for more informative, communicable, and well-rounded scientific reporting.

```{r}
pacman::p_load(readxl, performance, parameters, see)
```

### 3.1.2 Data Set

The dataset *ToyotaCorolla.xls* contains data on used cars on sale during the late summer of 2004 in the Netherlands. It has 1436 records containing attributes includingPrice, Age, Kilometers, Horsepower, and other specifications.

#### 3.1.2.1 Importing exam_data

In the code chunk below, read_csv() of readr package is used to import Exam_data.csv data file into R and save it as an tibble data frame called exam_data.

```{r}
car_resale <- read_xls("data/ToyotaCorolla.xls", 
                       "data")
car_resale

```

## 3.2 Visualising Models

### 3.2.1 **Multiple Regression Model using lm()**

The code chunk below is used to calibrate a multiple linear regression model by using *lm()* of Base Stats of R.

```{r}
model <- lm(Price ~ Age_08_04 + Mfg_Year + KM + 
              Weight + Guarantee_Period, data = car_resale)
model
```

### 3.2.2 **Model Diagnostic: checking for multicolinearity**

In the code chunk, [*check_collinearity()*](https://easystats.github.io/performance/reference/check_collinearity.html) of [**performance**](https://easystats.github.io/performance/) package.

```{r}
check_collinearity(model)
```

```{r}
check_c <- check_collinearity(model)
plot(check_c)
```

### 3.2.3 **Model Diagnostic: checking normality assumption**

In the code chunk, [*check_collinearity()*](https://easystats.github.io/performance/reference/check_collinearity.html) of [**performance**](https://easystats.github.io/performance/) package.

```{r}
model1 <- lm(Price ~ Age_08_04 + KM + 
              Weight + Guarantee_Period, data = car_resale)
```

```{r}
check_n <- check_normality(model1)
```

```{r}
plot(check_n)
```

### 3.2.4 **Model Diagnostic: Check model for homogeneity of variances**

In the code chunk, [*check_heteroscedasticity()*](https://easystats.github.io/performance/reference/check_heteroscedasticity.html) of [**performance**](https://easystats.github.io/performance/) package.

```{r}
check_h <- check_heteroscedasticity(model1)
```

```{r}
plot(check_h)
```

### 3.2.5 **Model Diagnostic: Complete check**

We can also perform the complete by using [*check_model()*](https://easystats.github.io/performance/reference/check_model.html).

```{r}
check_model(model1)
```

### 3.2.6 **Visualising Regression Parameters: see methods**

In the code below, `plot()` of see package and `parameters()` of parameters package is used to visualise the parameters of a regression model.

```{r}
plot(parameters(model1))
```

### 3.2.7 **Visualising Regression Parameters: ggcoefstats() methods**

In the code below, [*ggcoefstats()*](https://indrajeetpatil.github.io/ggstatsplot/reference/ggcoefstats.html) of ggstatsplot package to visualise the parameters of a regression model.

```{r}
ggcoefstats(model1, 
            output = "plot")
```
