---
title: "Hands-on Exercise 5"
subtitle: "Lesson 5: Visual Multivariate Analysis" 
author: "Victoria Neo"
date: 02/9/2024
date-modified: last-modified
format:
  html:
    code-fold: true
    code-summary: "code block"
    code-tools: true
    code-copy: true
execute:
  warning: false
---

![*Taken from* [xkcd A webcomic of romance, sarcasm, math, and language.](https://xkcd.com/523/)](images/clipboard-140603643.png){fig-align="left" width="403"}

# Overview Summary

|                  |                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                   |
|--------------------------|----------------------------------------------|
| Work done        | Hands-on Exercise 5                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                               |
| Hours taken      | ⏱️⏱️⏱️⏱️⏱️⏱️ ( hospitalisation leave)                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                             |
| Questions        | 0                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                 |
| How do I feel?   | 🥱                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                |
| What do I think? | Wow I came across funnel plots while doing my own research and there must be a distinction drawn between [funnel charts](https://chartio.com/learn/charts/funnel-chart-complete-guide/#:~:text=A%20funnel%20chart%20is%20a,funnel's%20width%20as%20it%20narrows.) (used to demonstrate the flow of users through a business or sales process) and [funnel plots](https://s4be.cochrane.org/blog/2023/06/27/how-to-read-a-funnel-plot/#:~:text=What%20is%20a%20funnel%20plot,and%20results%20of%20individual%20studies.) (are scatterplots that compares the precision (how close the estimated intervention effect size is to the true effect size) and results of individual studies. Of course, funnel plots are more exciting! |

# 1 [Building Ternary Plot with R](https://r4va.netlify.app/chap13)

Ternary plots are a way of displaying the distribution and variability of three-part compositional data. (For example, the proportion of aged, economy active and young population or sand, silt, and clay in soil.) It’s display is a triangle with sides scaled from 0 to 1. Each side represents one of the three components. A point is plotted so that a line drawn perpendicular from the point to each leg of the triangle intersect at the component values of the point.

In this hands-on, you will learn how to build ternary plot programmatically using R for visualising and analysing population structure of Singapore.

The hands-on exercise consists of four steps:

-   Install and launch **tidyverse** and **ggtern** packages.

-   Derive three new measures using *mutate()* function of **dplyr** package.

-   Build a static ternary plot using *ggtern()* function of **ggtern** package.

-   Build an interactive ternary plot using *plot-ly()* function of **Plotly R** package.

## 1.2: Data

### 1.2.1 Installing and loading the required libraries

For this exercise, two main R packages will be used in this hands-on exercise, they are:

-   [**ggtern**](http://www.ggtern.com/), a ggplot extension specially designed to plot ternary diagrams. The package will be used to plot static ternary plots.

-   [**Plotly R**](https://plot.ly/r/), an R package for creating interactive web-based graphs via plotly’s JavaScript graphing library, plotly.js . The **plotly R** libary contains the *ggplotly* function, which will convert **ggplot2** figures into a Plotly object.

We will also need to ensure that selected **tidyverse** family packages namely: **readr**, **dplyr** and **tidyr** are also installed and loaded.

In this exercise, version 3.2.1 of **ggplot2** will be installed instead of the latest version of **ggplot2**. This is because the current version of **ggtern** package is not compatible to the latest version of **ggplot2**.

```{r}
pacman::p_load(plotly, ggtern, tidyverse) 
```

### 1.2.2 Data Preparation

#### 1.2.2.1 **Data Set**

For the purpose of this hands-on exercise, the [Singapore Residents by Planning Area Subzone, Age Group, Sex and Type of Dwelling, June 2000-2018](https://www.singstat.gov.sg/find-data/search-by-theme/population/geographic-distribution/latest-data) data will be used. The data set has been downloaded and included in the data sub-folder of the hands-on exercise folder. It is called *respopagsex2000to2018_tidy.csv* and is in csv file format.

#### 1.2.2.2 **Importing Data**

To import *respopagsex2000to2018_tidy.csv* into R, ***read_csv()*** function of **readr** package will be used.

```{r}
#Reading the data into R environment
pop_data <- read_csv("data/respopagsex2000to2018_tidy.csv") 
```

#### 1.2.2.3 **Preparing the Data**

Next, use the mutate() function of dplyr package to derive three new measures, namely: young, active, and old.

```{r}
#Deriving the young, economy active and old measures
agpop_mutated <- pop_data %>%
  mutate(`Year` = as.character(Year))%>%
  spread(AG, Population) %>%
  mutate(YOUNG = rowSums(.[4:8]))%>%
  mutate(ACTIVE = rowSums(.[9:16]))  %>%
  mutate(OLD = rowSums(.[17:21])) %>%
  mutate(TOTAL = rowSums(.[22:24])) %>%
  filter(Year == 2018)%>%
  filter(TOTAL > 0)
```

## 1.3: Plotting Ternary Diagram with R

### 1.3.1 Plotting a static ternary diagram

Use ***ggtern()*** function of **ggtern** package to create a simple ternary plot.

::: panel-tabset
## Building the static ternary plot

```{r}
ggtern(data=agpop_mutated,aes(x=YOUNG,y=ACTIVE, z=OLD)) +
  geom_point()
```

## Building the static ternary plot with `theme_rgbw()`

```{r}
ggtern(data=agpop_mutated, aes(x=YOUNG,y=ACTIVE, z=OLD)) +
  geom_point() +
  labs(title="Population structure, 2015") +
  theme_rgbw()
```

## Building the static ternary plot with `theme_minimal()`

```{r}
ggtern(data=agpop_mutated, aes(x=YOUNG,y=ACTIVE, z=OLD)) +
  geom_point() +
  labs(title="Population structure, 2015") +
  theme_minimal()
```
:::

### 1.3.2 Plotting an interative ternary diagram

The code below create an interactive ternary plot using ***plot_ly()*** function of **Plotly R**.

::: panel-tabset
## Building the interactive ternary plot

```{r}
# reusable function for creating annotation object
label <- function(txt) {
  list(
    text = txt, 
    x = 0.1, y = 1,
    ax = 0, ay = 0,
    xref = "paper", yref = "paper", 
    align = "center",
    font = list(family = "serif", size = 15, color = "white"),
    bgcolor = "#b3b3b3", bordercolor = "black", borderwidth = 2
  )
}

# reusable function for axis formatting
axis <- function(txt) {
  list(
    title = txt, tickformat = ".0%", tickfont = list(size = 10)
  )
}

ternaryAxes <- list(
  aaxis = axis("Young"), 
  baxis = axis("Active"), 
  caxis = axis("Old")
)

# Initiating a plotly visualization 
plot_ly(
  agpop_mutated, 
  a = ~YOUNG, 
  b = ~ACTIVE, 
  c = ~OLD, 
  color = I("black"), 
  type = "scatterternary"
) %>%
  layout(
    annotations = label("Ternary Markers"), 
    ternary = ternaryAxes
  )
```
:::

# To see the rest of Hands-on Exercise 5, please refer to the following links:

[**Hands-on Exercise 5b: Visual Correlation Analysis**](/Hands-on_Ex/Hands-on_Ex05/Hands-on_Ex05b.html)
